\part{Other topics}
\section{Modeling friction}
Most models are ad-hoc and empirical. 
\subsection{Coloumb friction model}
\begin{align}
    F_f = -\mu F_N sgn(v)
\end{align}
Problem if we miss the exact time when v=0

\subsection{Viscous friction}
Has several models here, I will not go in detail.

\subsection{Static friction}
Karnopp + Coloumb model
\begin{align}
    F_f = \min(-u,F_N) \;\;\; if \; v=0 \\
    F_f = -\mu F_N sgn(v) \;\;\; if \; v\neq0
\end{align}

\section{Hybrid systems}
Combination of continuous states and discrete states.

\subsection{Guard conditions}
\begin{align}
    G(q_i,q_j): Q \times Q \rightarrow X_G\subseteq X
\end{align}
defines the set of X for which a jump from a discrete state $q_i$ to $q_j$ is triggered.

\subsection{Reset map}
\begin{align}
    \mathcal{R}(\bullet,\bullet,\bullet): Q \times Q \times X \rightarrow X_R \subseteq X
\end{align}
defines how the continuous states are affected via a jump from a discrete state $q_i$ to $q_j$. Example:
\begin{align}
    \mathcal{R}(q_1,q_2,x) = \frac{2}{3} x
\end{align}
typical example from car where a gear shift from gear $q_1$ to $q_2$ on the state $x=rpm$ reduces the rpm.

\section{Event-Based Integrators}
The method for properly simulating hybrid systems. An event is hitting a guard condition so that we get a jump in discrete state. Event conditions $e(x)=0$,$e(x)$ are positive on one side of guard and negative on the other. Integrator simply monitors $e(x)$ and if it changes sign the integrator reduces the step s.t. we hit the $e(x)=0$